### 2012-11-10 paper.tex
added in sections

\documentclass[]{article}
\renewcommand{\rmdefault}{ptm}
\usepackage[hmargin=30mm]{geometry}
\usepackage{tabularx}

\begin{document}

\begin{table}[here]
\centering
\begin{tabularx}{0.7\textwidth}{| X | X | }
\hline
Topic of Survey						& 	Cloud Computing\\
\hline
Length of Survey						& 	15 pages\\
\hline
First team member's matriculation number	&	U096833E\\
\hline
First team member's name				&	LAURENCE PUTRA FRANSLAY\\
\hline
Sections done by first member				& 	\\
\hline
Second team member's matriculation number	&	\\
\hline
Second team member's name				&	\\
\hline
Sections done by second member			& 	\\
\hline

\hline
\end{tabularx}
\end{table}

\pagebreak

\section{Introduction}

%9 pages
\section{Introduction to the various methods}
%fill in the method names as the subsection header
\subsection{MapReduce: Simplified Data Processing on Large Clusters}

\subsection{Camdoop}
%Preprocessing instead of post processing

\subsection{Mantri}
% Load balancing tasks and reduce the effects caused by outliers

\subsection{Nectar}

\subsection{Optimising data shuffling in Data Parallel Computation}

%3 pages
\section{Advantages and Disadvantages of each method}
%fill in the method names as the subsection header
\subsection{}

\subsection{}

\subsection{}

\subsection{}

\subsection{}

%1 page
\section{Creating a more robust combination}

%1 page
\section{Conclusion}

\end{document}